Prop Oc Data Chronic Wounds

### 2021-08-22 R/Prop Oc Data Clean.Rmd

#Add New variaible to the data frame capping the value pf propOc to 1
rawPropOcDataClean$PropOc1= rawPropOcDataClean$PropOc
rawPropOcDataClean$PropOc1[rawPropOcDataClean$PropOc1>1] = 1
rawPropOcDataClean$PropOc1[rawPropOcDataClean$PropOc1<0] = 0
#make new data frame for downstream analysis
PropOcDataFiltered<-rawPropOcDataClean %>% filter(PropOc<1.5) %>% filter(PropOc>=0)
#make column factors 
PropOcDataFiltered$SampleCode<-as.factor(PropOcDataFiltered$SampleCode)
PropOcDataFiltered$TimePoint<-as.factor(PropOcDataFiltered$TimePoint)
PropOcDataFiltered$Condition<-as.factor(PropOcDataFiltered$Condition)
PropOcDataFiltered$source<-as.factor(PropOcDataFiltered$source)
PropOcDataFiltered$Replicate<-as.factor(PropOcDataFiltered$Replicate)

#From this point on there are at least two ways to move forward. The simplest one is to get the mean, mode and variance in the data
meanPOData<-PropOcDataFiltered %>% group_by(SampleCode, TimePoint, source, Distance, Condition) %>% summarise(MeanPO = mean(PropOc1), MedianPO = median(PropOc1), SDPO = sd(PropOc1), SEPO = std.error(PropOc1))

@@

#Calculate the Enrichment
pOcEnriMean<- pOMeanMeta %>% group_by(SampleCode,Relation)%>% mutate(Enrichment = max(MeanPO)) %>%
  mutate(EnrichmentCh = ifelse(Enrichment == MeanPO, 1, 0 )) %>% filter(EnrichmentCh==1) %>% 
  mutate(DoubledValues = sum(EnrichmentCh)) %>% ungroup()#find max
###Sorry this is out of order
lowCountSamples<-bind_rows(pOcEnriMean %>% filter(SampleCode == "mono_d1_02"),
                           pOcEnriMean %>% filter(SampleCode == "phz_d4_08"),
                           pOcEnriMean %>% filter(SampleCode == "wt_d4_11") %>% filter(source == "finalRG"))
pOcEnriMean <- anti_join(pOcEnriMean, lowCountSamples)

PlotEnrichmentMean<-function(relation){
  ggplot()+
    geom_point(data = pOcEnriMean %>% filter(Relation == relation),aes(y=MeanPO,x = SampleCode,shape = Region,  color =Condition),size=2)+
    #geom_label(data = propOcEnrichment %>% filter(Relation == relation),aes(y=30, x = Sample, label = Value),size=2)+
    labs(title =paste("50 percentile Enrichment", relation))+
    facet_grid(.~TimePoint,scales = "free", space = "free")+
    #scale_color_gradientn(colours = rainbow(5))+
    #scale_shape_manual(values=c(0,1,2, 15,16,17))+
    theme_cowplot()+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),plot.title = element_text(hjust = 0.5))
}
PlotEnrichmentMean("Pa->Sa")
PlotEnrichmentMean("Sa->Pa")
PlotEnrichmentMean("Sa->Sa")
PlotEnrichmentMean("Pa->Pa")

#Get stats across samples
pOMeanStatsRgn<-pOMeanMeta %>% ungroup() %>% group_by(Condition,Culture,Strain,TimePoint, Distance, Relation, Region) %>% summarize(meanValue = mean(MeanPO), sdValue = sd(MeanPO), medianValue = median(MeanPO))

```

>>>>

#Add New variable to the data frame capping the value pf propOc to 1
rawPropOcDataClean$PropOc1= rawPropOcDataClean$PropOc
rawPropOcDataClean$PropOc1[rawPropOcDataClean$PropOc1>1] = 1
rawPropOcDataClean$PropOc1[rawPropOcDataClean$PropOc1<0] = 0
#make new data frame for downstream analysis
PropOcDataFiltered<-rawPropOcDataClean %>% filter(PropOc<1.5) %>% filter(PropOc>=0)
#make column factors 
PropOcDataFiltered$SampleCode<-as.factor(PropOcDataFiltered$SampleCode)
PropOcDataFiltered$TimePoint<-as.factor(PropOcDataFiltered$TimePoint)
PropOcDataFiltered$Condition<-as.factor(PropOcDataFiltered$Condition)
PropOcDataFiltered$source<-as.factor(PropOcDataFiltered$source)
PropOcDataFiltered$Replicate<-as.factor(PropOcDataFiltered$Replicate)

#From this point on there are at least two ways to move forward. The simplest one is to get the mean, mode and variance in the data
```

```{r meanPO}
meanPOData<-PropOcDataFiltered %>% group_by(SampleCode, TimePoint, source, Distance, Condition) %>% summarise(MeanPO = mean(PropOc1), MedianPO = median(PropOc1), SDPO = sd(PropOc1), SEPO = std.error(PropOc1))

@@

###Moving forward I could calculate the mean PO for all samples and then get the enrichment from there or calculate the enrichment and then get mean enrichment for all samples. I'll do the former first:

#---Mean, then enrichment
#####
#Get the mean meanPO across samples
##Group including Region
pOMeanStatsRgn<-pOMeanMeta %>% ungroup() %>% group_by(Condition,Culture,Strain,TimePoint, Distance, Relation, Region) %>% summarize(meanValue = mean(MeanPO), sdValue = sd(MeanPO), medianValue = median(MeanPO))
##Group without region
pOMeanStatsNoRgn<-pOMeanMeta %>% ungroup() %>% group_by(Condition,Culture,Strain,TimePoint, Distance, Relation) %>% summarize(meanValue = mean(MeanPO), sdValue = sd(MeanPO), medianValue = median(MeanPO))

PlotMeanMeanPO<-function(df,relation){
  ggplot()+
    #geom_point(data = df %>% filter(Relation == relation),aes(y=meanValue,x = Distance,shape = Region,  color =Condition),size=2)+
    #geom_pointrange(data = df %>% filter(Relation == relation),aes(y=meanValue,ymax=meanValue+sdValue,ymin=meanValue-sdValue,x = Distance,shape = Region,  color =Condition))+
    geom_pointrange(data = df %>% filter(Relation == relation),aes(y=meanValue,ymax=meanValue+sdValue,ymin=meanValue-sdValue,x = Distance,color =Culture, shape = Region), position = position_dodge(width = .5))+
    #geom_label(data = propOcEnrichment %>% filter(Relation == relation),aes(y=30, x = Sample, label = Value),size=2)+
    labs(title =paste("Mean MeanPO", relation))+
    facet_grid(Strain~TimePoint,scales = "free", space = "free")+
    #scale_color_gradientn(colours = rainbow(5))+
    #scale_shape_manual(values=c(0,1,2, 15,16,17))+
    theme_cowplot()+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),plot.title = element_text(hjust = 0.5))
}
PlotMeanMeanPO(pOMeanStatsRgn, "Pa->Sa")
PlotMeanMeanPO(pOMeanStatsNoRgn, "Pa->Sa")
PlotMeanMeanPO(pOMeanStatsRgn,"Sa->Pa")
PlotMeanMeanPO(pOMeanStatsNoRgn,"Sa->Pa")
PlotMeanMeanPO(pOMeanStatsRgn,"Sa->Sa")
PlotMeanMeanPO(pOMeanStatsNoRgn,"Sa->Sa")
PlotMeanMeanPO(pOMeanStatsRgn,"Pa->Pa")
PlotMeanMeanPO(pOMeanStatsNoRgn,"Pa->Pa")

pOMeanEnri<- pOMeanStatsRgn %>% group_by(Condition,Culture, Strain, TimePoint,Relation, Region)%>% 
  mutate(Enrichment = max(meanValue)) %>%
  mutate(EnrichmentCh = ifelse(Enrichment == meanValue, 1, 0 )) %>% filter(EnrichmentCh==1) %>% 
  mutate(DoubledValues = sum(EnrichmentCh)) %>% ungroup()#find max

#PlotMeanEnrichment<-function(relation){
  ggplot()+
    geom_point(data = pOMeanEnri,aes(y=Distance,x = Condition,color = Region),position = position_dodge(width = .5),size=2)+
    #geom_label(data = propOcEnrichment %>% filter(Relation == relation),aes(y=30, x = Sample, label = Value),size=2)+
    labs(title =paste("Enrichment from mean PO"))+
    facet_grid(TimePoint~Relation,scales = "free", space = "free")+
    #scale_color_gradientn(colours = rainbow(5))+
    #scale_shape_manual(values=c(0,1,2, 15,16,17))+
    theme_cowplot()+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),plot.title = element_text(hjust = 0.5))+
    panel_border() 
#}

#####
#---Enrichment, then mean
#####
  #Calculate the Enrichment for all samples
pOcEnriMean<- pOMeanMeta %>% group_by(SampleCode,Relation)%>% mutate(Enrichment = max(MeanPO)) %>%
  mutate(EnrichmentCh = ifelse(Enrichment == MeanPO, 1, 0 )) %>% filter(EnrichmentCh==1) %>% 
  mutate(DoubledValues = sum(EnrichmentCh)) %>% ungroup()#find max
###Sorry this is out of order
lowCountSamples<-bind_rows(pOcEnriMean %>% filter(SampleCode == "mono_d1_02"),
                           pOcEnriMean %>% filter(SampleCode == "phz_d4_08"),
                           pOcEnriMean %>% filter(SampleCode == "wt_d4_11") %>% filter(source == "finalRG"))
pOcEnriMean <- anti_join(pOcEnriMean, lowCountSamples)

PlotEnrichmentMean<-function(relation){
  ggplot()+
    geom_point(data = pOcEnriMean %>% filter(Relation == relation),aes(y=MeanPO,x = SampleCode,shape = Region,  color =Condition),size=2)+
    #geom_label(data = propOcEnrichment %>% filter(Relation == relation),aes(y=30, x = Sample, label = Value),size=2)+
    labs(title =paste("50 percentile Enrichment", relation))+
    facet_grid(.~TimePoint,scales = "free", space = "free")+
    #scale_color_gradientn(colours = rainbow(5))+
    #scale_shape_manual(values=c(0,1,2, 15,16,17))+
    theme_cowplot()+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),plot.title = element_text(hjust = 0.5))
}
PlotEnrichmentMean("Pa->Sa")
PlotEnrichmentMean("Sa->Pa")
PlotEnrichmentMean("Sa->Sa")
PlotEnrichmentMean("Pa->Pa")
#####
```
